\pkg{Rgbp}: An \proglang{R} Package for Conjugate Gaussian, Poisson, and Binomial Hierarchical Modeling and Frequency Method Checking on Overdispersed Data

### 2015-06-10 jss/article_v2.tex
draft done

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Hyungsuk Tak, Joseph Kelly, Carl Morris} %% comma-separated
\Plaintitle{Rgbp: Hierarchical Modeling and Frequency Method Checking} %% without formatting
\Shorttitle{\pkg{Rgbp}: Hierarchical Modeling and Frequency Method Checking on Overdispersed Data} %% a short title (if necessary)
%% an abstract and keywords
\Abstract{
\pkg{Rgbp} is an \proglang{R} package that utilizes  approximate Bayesian machinery to fit two-level conjugate hierarchical models on overdispersed Gaussian, Poisson, and Binomial data. The data that \pkg{Rgbp} assumes comprise of observed sufficient statistics for each random effect, such as averages or proportions, possibly together with covariates of each group but without population-level data. The approximate Bayesian tool equipped with the adjustment for density maximization produces point and interval estimates for each random effect, point estimates and standard errors for regression coefficients, and a point estimate and its standard error for a second-level variance component. For the Binomial data, the package provides an option to produce  posterior samples of all the model parameters via the acceptance-rejection method.  The main goal of \pkg{Rgb} is to produce approximate Bayesian interval estimates for the random effects that meet their nominal confidence levels, and the package uses unique improper hyper-prior distributions for that purpose.  \pkg{Rgbp} provides a quick way to check whether the resultant Bayesian interval estimates for the random effects achieve the nominal confidence levels via a repeated sampling coverage evaluation, which we call  ``frequency method checking."  }
% and evaluates repeated sampling frequency property of the resulting approximate Bayesian interval estimates for random effects
%It is found that such Bayesian-frequentist reconciliation allows \pkg{Rgbp} to have attributes desirable from both perspectives, working well in small samples and yielding good coverage probabilities for its interval estimates.
\Keywords{overdispersion, hierarchical model, adjustment for density maximization, repeated sampling coverage evaluation, \proglang{R}}
\Plainkeywords{multilevel model, conjugate hierarchical generalized linear models, frequency method checking, coverage probability, shrinkage, R} 
%% without formatting
%% at least one keyword must be supplied

@@

In addition to fitting hierarchical models, \pkg{Rgbp} provides a quick way to evaluate the repeated sampling coverage rates of the resulting Bayesian interval estimates for random effects \citep{morris1997, daniels1999prior, tang2002fitting, tang2011, morris2012}. It is a unique procedure that distinguishes \pkg{Rgbp}  from any other \proglang{R} packages for hierarchical modeling such as \pkg{hglm} \citep{hglm2010, ronnegaard2011hglm} for conjugate hierarchical generalized models and \pkg{arm} \citep{gelman2014arm} for Bayesian hierarchical regression models. The evaluation procedure which we call ``frequency method checking'' adopts a parametric bootstrapping that generates mock data sets given the values of the hyper-parameters and estimates the coverage rates based on the generated mock data sets.

@@

One of the functions in \pkg{Rgbp}, \code{gbp}, fits a conjugate hierarchical model whose first-level has distributions of observed data and whose second-level has conjugate distributions on the first-level mean parameters (random effects). The \code{gbp} function allows users to choose one of three types of hierarchical models according to the type of data, namely Normal-Normal, Poisson-Gamma, and Binomial-Beta models.

\subsection[Normal-Normal]{Normal-Normal (``g''=Gaussian data)}
The following is the general Normal-Normal hierarchical model assumed by \code{gbp}. For reference,  $V_{j}~(\equiv \sigma^{2}/n_{j})$ below is assumed to be known, and subscript \emph{j} indicates the \emph{j}-th group among \emph{k} groups in the dataset. For $j=1, 2, \ldots, k$, 
\begin{eqnarray}
y_{j}\vert \mu_{j} & \stackrel{indep.}{\sim} &\textrm{Normal}(\mu_{j}, V_{j}),\label{normalobs}\\
\mu_{j}\vert \boldsymbol{\beta}, A & \stackrel{indep.}{\sim} &\textrm{Normal}(\mu^E_{j}, A),\label{normalprior}
\end{eqnarray}

where $\mu^E_{j} = E(\mu_j\vert \boldsymbol{\beta}, A) = \boldsymbol{x}_j^\top \boldsymbol{\beta}=\beta_{1}x_{j, 1}+\beta_{2}x_{j, 2} + \cdots + \beta_{m}x_{j, m}$ is the expected random effect and $m$ is the number of regression coefficients to be estimated. The default of the function \code{gbp} is to set $x_{j, 1}$ to 1 for an intercept term, though \code{gbp} also provides a usage without the intercept term.  It is assumed that the second-level variance $A$ is unknown and that the $m\times1$ regression coefficient vector $\boldsymbol{\beta}$ is also unknown unless otherwise specified. If no covariates are available with an intercept term, then $\boldsymbol{x}_j^\top \boldsymbol{\beta}=\beta_1$ ($m=1$) and thus $\mu^E_{j}=\mu^E=\beta_1$ for all $j$, resulting in an exchangeable conjugate prior distribution for the random effects.  Based on these conjugate prior distributions for random effects, it is easy to derive the conditional posterior distributions of the random effects. For $j=1, 2, \ldots, k$,

@@

\subsection[Poisson-Gamma]{Poisson-Gamma (``p''=Poisson data)}\label{poissonsubsec}
The function \code{gbp} is also capable of estimating a conjugate Poisson-Gamma hierarchical model, though its usage is limited to the case where the expected random effects are known ($m=0$).  For $j=1, 2, \ldots, k$, 
\begin{eqnarray}
y_j\vert \lambda_j &\stackrel{indep.}{\sim} & \textrm{Poisson}(n_{j}\lambda_{j}),\\
\lambda_{j}\vert r &\stackrel{i.i.d.}{\sim}& \textrm{Gamma}(r\lambda^E, r),
\end{eqnarray}
where $n_{j}$ is the exposure of group $j$, which is not necessarily an integer, $\lambda^E=E(\lambda_j\vert r)$ is the known expected random effect, and $r$ is the unknown second-level variance component. The mean and variance of the conjugate Gamma prior distribution are $\lambda^E$ and $\lambda^E/r$, respectively. We interpret $r$ as the amount of prior information as $n_{j}$ represents the amount of observed information, which makes intuitive sense because the uncertainty of the conjugate prior distribution increases as $r$ decreases; in the limit of $r$ going to 0, the conjugate prior distribution gets flatter. The conditional posterior distribution of the random effect $\lambda_j$ for this Poisson-Gamma model is
\begin{equation} \label{gammapost}
\lambda_j\vert r, \boldsymbol{y} \stackrel{indep.}{\sim}\textrm{Gamma}(r\lambda^E + n_j\bar{y}_{j},~ r + n_j),
\end{equation}  
whose $\bar{y}_j=y_j/n_j$. The mean and variance of the conditional posterior distribution are
\begin{equation}\label{gammapost_mean_var}
\lambda^\ast_j\equiv E(\lambda_j\vert r, \boldsymbol{y} )=(1-B_{j})\bar{y}_{j} + B_{j}\lambda^E~~\textrm{and}~~Var(\lambda_j\vert r, \boldsymbol{y})=\frac{\lambda^\ast_j}{r+n_j}.
\end{equation}
where $B_{j}\equiv r / (r+n_{j})$ for $j=1, 2, \ldots, k$.  The conditional posterior mean is  a convex combination of the observed sufficient statistic $\bar{y}_{j} = y_j/n_j$ and the expected random effect  $\lambda^E$ weighted by the relative amount of information in the prior compared to the data, called a shrinkage factor $B_j=r / (r + n_j)$. If the conjugate prior distribution contains more information than the observed data have, \emph{i.e.}, ensemble sample size $r$ exceeds individual sample size $n_{j}$, then the posterior mean shrinks  towards the prior mean by more than 50\%, borrowing more information from the prior distribution.

Note that the conditional posterior variance in Equation \ref{gammapost_mean_var} is linear in the conditional posterior mean, whereas a slightly different Poisson-Gamma model specification has been used elsewhere \citep{morris1997} that makes the variances in the prior and conditional posterior distributions  quadratic functions of their means respectively.

@@

The likelihood function of hyper-parameters $A$ and $\boldsymbol{\beta}$ for the Gaussian hierarchical model is derived from the independent Normal marginal distributions of the observed data with random effects, $\mu_1, \mu_2, \ldots, \mu_k$, integrated out as follows; %For reference, the marginal distribution of the observed datum for the Normal model  still follows a Normal distribution and that for the Poisson model follows a Negative-Binomial distribution.
\begin{equation}\label{marginal_normal}
L_g(A, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k f(y_j\vert A, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k \frac{1}{\sqrt{2\pi (A+V_j)}}\exp\big(-\frac{(y_j-\mu^E_j)^2}{2(A+V_j)}\big),
\end{equation}
where $\mu^E_j=\boldsymbol{x}^\top\boldsymbol{\beta}$. The likelihood function of hyper-parameters $r$ and $\boldsymbol{\beta}$ for the Binomial hierarchical model is derived from the independent Beta-Binomial marginal distributions of the observed data with random effects, $p_1, p_2, \ldots, p_k$, integrated out \citep{skellam1948}; %For reference, the marginal distribution of the observed datum for the Normal model  still follows a Normal distribution and that for the Poisson model follows a Negative-Binomial distribution.
\begin{equation}\label{marginal_binomial}
L_b(r, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k f(y_j\vert r, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k\binom{n_j}{y_j}\frac{B(y_j+rp^E_j, ~n_j-y_j+r(1-p^E_j))}{B(rp^E_j, ~r(1-p^E_j))},
\end{equation}
where $p^E_j=\exp(\boldsymbol{x}^\top\boldsymbol{\beta})/(1+\exp(\boldsymbol{x}^\top\boldsymbol{\beta}))$ and the notation $B(a, b)~(\equiv\int_0^1 v^{a-1}(1-v)^{b-1}dv)$ indicates a beta function for positive constants $a$ and $b$. Similarly, the likelihood function of $r$ for the Poisson hierarchical model, without a regression coefficient (a known expected random effect, $\lambda^E$), comes from the independent Negative-Binomial marginal distributions of the observed data with the random effects, $\lambda_1, \lambda_2, \ldots, \lambda_k$, integrated out;
\begin{equation}\label{marginal_poisson}
L_p(r)=\prod_{j=1}^k f(y_j\vert r)=\prod^{k}_{j=1} \frac{\Gamma(r \lambda^E+y_j)}{\Gamma(r\lambda^E)(y_j!)}(1-B_{j})^{y_{i}}B_{j}^{r \lambda^E},
\end{equation}
where $\Gamma(a)$ is a gamma function defined as $\int_0^\infty x^{a-1}e^{x}dx$ for a positive constant $a$.

@@

The joint posterior density  function of hyper-parameters $f(A, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})$ for the Gaussian hierarchical model is proportional to their likelihood functions in Equation \ref{marginal_normal} because we use flat improper hyper-prior density functions for $A$ and $\boldsymbol{\beta}$ shown in Equation \ref{eq:hyper};
\begin{equation}\label{marginal_post_normal}
f(A, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})\propto L_g(A, \boldsymbol{\boldsymbol{\beta}})d\boldsymbol{\beta}dA.
\end{equation}
The joint posterior density  function of hyper-parameters $f(r, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})$ for the Binomial hierarchical model is proportional to their likelihood functions in Equation \ref{marginal_binomial} multiplied by the hyper-prior density functions of $r$ and $\boldsymbol{\beta}$ in Equation \ref{eq:hyper} as follows;
\begin{equation}\label{marginal_post_binomial}
f(r, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})\propto L_b(r, \boldsymbol{\boldsymbol{\beta}})d\boldsymbol{\beta}dr/r^2.
\end{equation}
The posterior density function of $r$, $f(r\vert \boldsymbol{y})$, for the Poisson hierarchical model is  the likelihood function in Equation \ref{marginal_poisson} times the hyper-prior density function of $r$, $dr/r^2$;
\begin{equation}\label{marginal_post_poisson}
f(r \vert \boldsymbol{y})\propto L_p(r)dr/r^2.
\end{equation}

@@

\subsection{Estimation for random effects}
It is intractable to derive the unconditional posterior distribution of each random effect analytically for all three models. Instead, we  assume again that each random effect has a skewed Normal distribution \cite{azzalini1985class} for the Gaussian model, a Beta distribution for the Binomial model, or a Gamma distribution for the Poisson model; for $j=1, 2, \ldots, k$,
\begin{eqnarray}
\mu_j\vert\boldsymbol{y} &\sim& \textrm{skewed-Normal}(\phi, \omega, \delta),\label{skewnormal}\\
p_j\vert \boldsymbol{y} &\sim& \textrm{Beta}(t_{1j}, t_{0j}),\label{betapost}\\
\lambda_j\vert \boldsymbol{y} &\sim& \textrm{Gamma}(s_{1j}, s_{0j}).\label{gammapost}
\end{eqnarray}

\cite{kelly2014advances}  matches estimated three unconditional posterior  moments of the random effects to the three parameters $(\phi, \omega, \delta$) of the skewed-Normal distribution in Equation \ref{skewnormal}.

@@

Specifically, we assume
\begin{eqnarray}\label{env}
g_1(\alpha) &=& g_1(\alpha; \mu, \sigma, a, b)~\equiv~\textrm{Skewed-}t(\alpha\vert\mu, \sigma, a, b),\\
g_2(\boldsymbol{\beta}) &=& g_2(\boldsymbol{\beta}; \boldsymbol{\mu}^\ast, S_{(m\times m)})~\equiv~t_{4}(\boldsymbol{\beta}\vert \boldsymbol{\mu}^\ast, S),\label{g2approx}
\end{eqnarray}
where the notation Skewed-$t(\alpha\vert\mu, \sigma, a, b)$ represents a density function of a skewed $t$-distribution  at $\alpha$ with location $\mu$, scale $\sigma$, degree of freedom $a+b$, and skewness $a-b$ for any positive constants $a$ and $b$ \citep{jones2003skew}. The article of \cite{jones2003skew} derives the mode of $g_1(\alpha)$ as
\begin{equation}\label{mode}
\mu+\frac{(a-b)\sqrt{a+b}}{\sqrt{(2a+1)(2b+1)}}.
\end{equation} 
It shows that the tails follow a power law as $\alpha^{-(2a+1)}$ on the left and $\alpha^{-(2b+1)}$ on the right when $b>a$. It also provides a representation to generate the random variable following their skewed-$t$ distribution as
\begin{equation}
\alpha\sim \mu+\sigma\frac{\sqrt{a+b}(2T-1)}{2\sqrt{T(1-T)}}, ~\textrm{where}~ T\sim \textrm{Beta}(a, b).
\end{equation}
The notation $t_{4}(\boldsymbol{\beta}\vert\boldsymbol{\mu}^\ast, S)$ in Equation \ref{g2approx} indicates a density function of a multivariate $t$-distribution at $\boldsymbol{\beta}$ with 4 degrees of freedom, a location vector $\boldsymbol{\mu}^\ast$, and a $m\times m$ scale matrix $S$ that leads to the variance-covariance matrix $2S$.

@@

\subsection{Model fitting}
The function \code{gbp} creates an G3 object ``gbp'' which provides three relevant functions \code{plot}, \code{print}, and \code{summary}. To use these relevant functions, we need to save the outcome of \code{gbp} into an object (see the code below).

@@

The function \code{gbp} contains several optional arguments. The argument \code{Alpha}, whose default is 0.95, sets the confidence level, producing $100\times$\code{Alpha}\% interval estimates of the random effects. For the Gaussian model, setting the argument \code{normal.CI} to \code{TRUE} makes \code{gbp} use a Normal approximation to the unconditional posterior distribution of the random effect. The default of \code{normal.CI} is \code{FALSE} for the skewed Normal approximation.

@@

The G3 object ``gbp'', for example \code{b.output} above, contains detailed outcomes. The function \code{names} will show the list of outcomes (too many to be listed in this article); see the document of \code{gbp} to check the details of the items.
\begin{CodeChunk}
\begin{CodeInput}
R> names(b.output)
R> ?gbp
\end{CodeInput}
\end{CodeChunk}

The G3 object makes use of some functions, \code{print}, \code{summary}, and \code{plot}. The estimation result for all the random effects pops up if we type the ``gbp'' object at the \proglang{R} console, which is the same as the following function \code{print} with its default argument ``\code{sort = TRUE}'' that prints out the estimation result for all the groups in the ascending order of $n$ for the Binomial and Poisson model and the ascending order of standard errors for the Gaussian model. When the argument \code{sort} is \code{FALSE}, the estimation result comes out as the order of data input.
\begin{CodeChunk}
\begin{CodeInput}
R> b.output
R> print(b.output, sort = FALSE)
\end{CodeInput}
\end{CodeChunk}

>>>>

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Hyungsuk Tak, Joseph Kelly, Carl Morris} %% comma-separated
\Plaintitle{Rgbp: Hierarchical Modeling and Frequency Method Checking} %% without formatting
\Shorttitle{\pkg{Rgbp}: Hierarchical Modeling and Frequency Method Checking on Overdispersed Data} %% a short title (if necessary)
%% an abstract and keywords
\Abstract{
\pkg{Rgbp} is an \proglang{R} package that utilizes  approximate Bayesian machinery to fit two-level conjugate hierarchical models on overdispersed Gaussian, Poisson, and Binomial data. The data that \pkg{Rgbp} assumes comprise of observed sufficient statistics for each random effect, such as averages or proportions, possibly together with covariates of each group but without population-level data. The approximate Bayesian tool equipped with the adjustment for density maximization produces point and interval estimates for each random effect, point estimates and standard errors for regression coefficients, and a point estimate and its standard error for a second-level variance component. For the Binomial data, the package provides an option to produce  posterior samples of all the model parameters via the acceptance-rejection method.  The main goal of \pkg{Rgb} is to produce approximate Bayesian interval estimates for the random effects that meet their nominal confidence levels.   For this purpose, our models adopt unique improper hyper-prior distributions and \pkg{Rgbp} provides a quick way to check whether the resultant Bayesian interval estimates for the random effects achieve the nominal confidence levels via a repeated sampling coverage evaluation, which we call  ``frequency method checking."  }
% and evaluates repeated sampling frequency property of the resulting approximate Bayesian interval estimates for random effects
%It is found that such Bayesian-frequentist reconciliation allows \pkg{Rgbp} to have attributes desirable from both perspectives, working well in small samples and yielding good coverage probabilities for its interval estimates.
\Keywords{overdispersion, hierarchical model, adjustment for density maximization, repeated sampling coverage evaluation, \proglang{R}}
\Plainkeywords{multilevel model, conjugate hierarchical generalized linear models, frequency method checking, coverage probability, shrinkage, R} 
%% without formatting
%% at least one keyword must be supplied

@@

In addition to fitting hierarchical models, \pkg{Rgbp} provides a quick way to evaluate the repeated sampling coverage rates of the resulting Bayesian interval estimates for random effects \citep{morris1997, daniels1999prior, tang2002fitting, tang2011, morris2012}. It is a unique procedure that distinguishes \pkg{Rgbp}  from any other \proglang{R} packages for hierarchical modeling such as \pkg{hglm} \citep{hglm2010, ronnegaard2011hglm} for conjugate hierarchical generalized models and \pkg{arm} \citep{gelman2014arm} for Bayesian hierarchical regression models. The evaluation procedure which we call ``frequency method checking'' adopts a parametric bootstrapping, a Monte Carlo simulation, that generates mock data sets given the values of the hyper-parameters and estimates the coverage rates based on the generated mock data sets.

@@

One of the functions in \pkg{Rgbp}, \code{gbp}, fits a conjugate hierarchical model in which the first-level has distributions of observed data and the second-level has conjugate distributions on the first-level mean parameters (random effects) without any observed data on this level. Specifically, the observed data of the first-level are sufficient statistics, sample mean or proportions, for random effects with an assumption of the homogeneity within each group. These data appear often in a biological analysis on litter data, a meta analysis on independent studies, or small area estimation problems. The \code{gbp} function allows users to choose one of three types of hierarchical models according to the type of data, namely Normal-Normal, Poisson-Gamma, and Binomial-Beta models. There are possibly more hierarchical models depending on the types of data such as Exponential-Gamma \citep{tang2002fitting}, but we choose the three models because these are the most commonly-used models.

In this section, we specify the modeling details. Our motivation for the parametrization is to provide an intuitive shrinkage interpretation in inference and to facilitate the estimation procedure, considering that the shrinkage factors under our parametrization are a function of only a second-level variance component.

\subsection[Normal-Normal]{Normal-Normal (``g''=Gaussian data)}
The following is the general Normal-Normal hierarchical model assumed by \code{gbp}. The subscript \emph{j} below indicates the \emph{j}-th group among \emph{k} groups in the dataset. For $j=1, 2, \ldots, k$, 
\begin{eqnarray}
y_{j}\vert \mu_{j} & \stackrel{indep.}{\sim} &\textrm{Normal}(\mu_{j}, V_{j}),\label{normalobs}\\
\mu_{j}\vert \boldsymbol{\beta}, A & \stackrel{indep.}{\sim} &\textrm{Normal}(\mu^E_{j}, A),\label{normalprior}
\end{eqnarray}

where $y_j$ is an observed unbiased estimate (sample mean) for random effect $j$, $V_{j}~(\equiv \sigma^{2}/n_{j})$ is a completely known standard error of $y_j$, $\mu^E_{j}$ is an expected random effect defined as $E(\mu_j\vert \boldsymbol{\beta}, A) = \boldsymbol{x}_j^\top \boldsymbol{\beta}=\beta_{1}x_{j, 1}+\beta_{2}x_{j, 2} + \cdots + \beta_{m}x_{j, m}$, and $m$ is the number of regression coefficients to be estimated. The default of the function \code{gbp} is to set $x_{j, 1}$ to 1 for an intercept term, though \code{gbp} also provides a usage without the intercept term.  It is assumed that the second-level variance $A$ is unknown and that the $m\times1$ regression coefficient vector $\boldsymbol{\beta}$ is also unknown unless otherwise specified. If no covariates are available with an intercept term, then $\boldsymbol{x}_j^\top \boldsymbol{\beta}=\beta_1$ ($m=1$) and thus $\mu^E_{j}=\mu^E=\beta_1$ for all $j$, resulting in an exchangeable conjugate prior distribution for the random effects.  Based on these conjugate prior distributions for random effects, it is easy to derive the conditional posterior distributions of the random effects. For $j=1, 2, \ldots, k$,

@@

\subsection[Poisson-Gamma]{Poisson-Gamma (``p''=Poisson data)}\label{poissonsubsec}
The function \code{gbp} is also capable of estimating a conjugate Poisson-Gamma hierarchical model, though its usage is limited to the case where the expected random effects are known ($m=0$), for example, we may have some information about the mean of the population (second-level) from which we collect the first-level group data.  For $j=1, 2, \ldots, k$, 
\begin{eqnarray}
y_j\vert \lambda_j &\stackrel{indep.}{\sim} & \textrm{Poisson}(n_{j}\lambda_{j}),\\
\lambda_{j}\vert r &\stackrel{i.i.d.}{\sim}& \textrm{Gamma}(r\lambda^E, r),
\end{eqnarray}
where $n_{j}$ is the exposure of group $j$, which is not necessarily an integer, $\lambda^E=E(\lambda_j\vert r)$ is the known expected random effect, and $r$ is the unknown second-level variance component. The mean and variance of the conjugate Gamma prior distribution are $\lambda^E$ and $\lambda^E/r$, respectively. We interpret $r$ as the amount of prior information as $n_{j}$ represents the amount of observed information, which makes intuitive sense because the uncertainty of the conjugate prior distribution increases as $r$ decreases; in the limit of $r$ going to 0, the conjugate prior distribution gets flatter. The conditional posterior distribution of the random effect $\lambda_j$ for this Poisson-Gamma model is
\begin{equation} \label{gammapost}
\lambda_j\vert r, \boldsymbol{y} \stackrel{indep.}{\sim}\textrm{Gamma}(r\lambda^E + n_j\bar{y}_{j},~ r + n_j),
\end{equation}  
whose $\bar{y}_j=y_j/n_j$. The mean and variance of the conditional posterior distribution are
\begin{equation}\label{gammapost_mean_var}
\lambda^\ast_j\equiv E(\lambda_j\vert r, \boldsymbol{y} )=(1-B_{j})\bar{y}_{j} + B_{j}\lambda^E~~\textrm{and}~~Var(\lambda_j\vert r, \boldsymbol{y})=\frac{\lambda^\ast_j}{r+n_j}.
\end{equation}
where $B_{j}\equiv r / (r+n_{j})$ for $j=1, 2, \ldots, k$.  The conditional posterior mean is  a convex combination of the observed sufficient statistic $\bar{y}_{j} = y_j/n_j$ and the expected random effect  $\lambda^E$ weighted by the relative amount of information in the prior compared to the data, called a shrinkage factor $B_j=r / (r + n_j)$. If the conjugate prior distribution contains more information than the observed data have, \emph{i.e.}, ensemble sample size $r$ exceeds individual sample size $n_{j}$, then the posterior mean shrinks  towards the prior mean by more than 50\%, borrowing more information from the prior distribution.

Note that the conditional posterior variance in Equation \ref{gammapost_mean_var} is linear in the conditional posterior mean, whereas a slightly different Poisson-Gamma model specification has been used elsewhere \citep{morris1997} that makes the variances in the prior and conditional posterior distributions  quadratic functions of their means respectively. The parametrization of the Poisson-Gamma model adopted in \pkg{Rgbp} makes shrinkage factors as a function of only the hyper-parameter $r$, while the shrinkage factors based on the parametrization in \cite{morris1997} are functions of $r$ and $\boldsymbol{\beta}$. Thus, the estimation procedure and the interpretation of the shrinkage factors are more intuitive under the parametrization in \pkg{Rgbp}.

@@

The likelihood function of hyper-parameters $A$ and $\boldsymbol{\beta}$ for the Gaussian hierarchical model is derived from the independent Normal marginal distributions of the observed data with random effects, $\mu_1, \mu_2, \ldots, \mu_k$, integrated out as follows; %For reference, the marginal distribution of the observed datum for the Normal model  still follows a Normal distribution and that for the Poisson model follows a Negative-Binomial distribution.
\begin{equation}\label{marginal_normal}
L_g(A, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k f(y_j\vert A, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k \frac{1}{\sqrt{2\pi (A+V_j)}}\exp\big(-\frac{(y_j-\mu^E_j)^2}{2(A+V_j)}\big),
\end{equation}
where $\mu^E_j=\boldsymbol{x}^\top\boldsymbol{\beta}$. The joint posterior density  function of hyper-parameters $f(A, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})$ for the Gaussian hierarchical model is proportional to their likelihood functions in Equation \ref{marginal_normal} because we use flat improper hyper-prior density functions for $A$ and $\boldsymbol{\beta}$ in Equation \ref{eq:hyper};
\begin{equation}\label{marginal_post_normal}
f(A, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})\propto L_g(A, \boldsymbol{\boldsymbol{\beta}})d\boldsymbol{\beta}dA.
\end{equation}

The likelihood function of hyper-parameters $r$ and $\boldsymbol{\beta}$ for the Binomial hierarchical model is derived from the independent Beta-Binomial marginal distributions of the observed data with random effects, $p_1, p_2, \ldots, p_k$, integrated out \citep{skellam1948}; %For reference, the marginal distribution of the observed datum for the Normal model  still follows a Normal distribution and that for the Poisson model follows a Negative-Binomial distribution.
\begin{equation}\label{marginal_binomial}
L_b(r, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k f(y_j\vert r, \boldsymbol{\boldsymbol{\beta}})=\prod_{j=1}^k\binom{n_j}{y_j}\frac{B(y_j+rp^E_j, ~n_j-y_j+r(1-p^E_j))}{B(rp^E_j, ~r(1-p^E_j))},
\end{equation}
where $p^E_j=\exp(\boldsymbol{x}^\top\boldsymbol{\beta})/(1+\exp(\boldsymbol{x}^\top\boldsymbol{\beta}))$ and the notation $B(a, b)~(\equiv\int_0^1 v^{a-1}(1-v)^{b-1}dv)$ indicates a beta function for positive constants $a$ and $b$. The joint posterior density  function of hyper-parameters $f(r, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})$ for the Binomial hierarchical model is proportional to their likelihood functions in Equation \ref{marginal_binomial} multiplied by the hyper-prior density functions of $r$ and $\boldsymbol{\beta}$ in Equation \ref{eq:hyper} as follows;
\begin{equation}\label{marginal_post_binomial}
f(r, \boldsymbol{\boldsymbol{\beta}}\vert \boldsymbol{y})\propto L_b(r, \boldsymbol{\boldsymbol{\beta}})d\boldsymbol{\beta}dr/r^2.
\end{equation}

Similarly, the likelihood function of $r$ for the Poisson hierarchical model, without a regression coefficient (a known expected random effect, $\lambda^E$), comes from the independent Negative-Binomial marginal distributions of the observed data with the random effects, $\lambda_1, \lambda_2, \ldots, \lambda_k$, integrated out;
\begin{equation}\label{marginal_poisson}
L_p(r)=\prod_{j=1}^k f(y_j\vert r)=\prod^{k}_{j=1} \frac{\Gamma(r \lambda^E+y_j)}{\Gamma(r\lambda^E)(y_j!)}(1-B_{j})^{y_{i}}B_{j}^{r \lambda^E},
\end{equation}
where $\Gamma(a)$ is a gamma function defined as $\int_0^\infty x^{a-1}e^{x}dx$ for a positive constant $a$. The posterior density function of $r$, $f(r\vert \boldsymbol{y})$, for the Poisson hierarchical model is  the likelihood function in Equation \ref{marginal_poisson} times the hyper-prior density function of $r$, $dr/r^2$;
\begin{equation}\label{marginal_post_poisson}
f(r \vert \boldsymbol{y})\propto L_p(r)dr/r^2.
\end{equation}

@@

@@

\subsection{Estimation for random effects}
It is intractable to derive analytically the unconditional posterior distribution of each random effect  for our three models, which motivate us to assume again that each random effect has a skewed Normal distribution \cite{azzalini1985class} for the Gaussian model, a Beta distribution for the Binomial model, or a Gamma distribution for the Poisson model; for $j=1, 2, \ldots, k$,
\begin{eqnarray}
\mu_j\vert\boldsymbol{y} &\sim& \textrm{skewed-Normal}(\phi, \omega, \delta),\label{skewnormal}\\
p_j\vert \boldsymbol{y} &\sim& \textrm{Beta}(t_{1j}, t_{0j}),\label{betapost}\\
\lambda_j\vert \boldsymbol{y} &\sim& \textrm{Gamma}(s_{1j}, s_{0j}).\label{gammapost}
\end{eqnarray}

\cite{kelly2014advances}  matches the estimated three unconditional posterior  moments of the random effects to the three parameters $(\phi, \omega, \delta$) of the skewed-Normal distribution in Equation \ref{skewnormal}, i.e., location parameter $\phi$, scale parameter $\omega$, and a skewness parameter $\delta$. \cite{kelly2014advances} also shows that the skewed-Normal approximation improves the Normal approximation \citep{tang2011, morris2012} in terms of the repeated sampling coverage rates of random effects.

@@

Specifically, we assume
\begin{eqnarray}\label{env}
g_1(\alpha) &=& g_1(\alpha; \mu, \sigma, a, b)~\equiv~\textrm{Skewed-}t(\alpha\vert\mu, \sigma, a, b),\\
g_2(\boldsymbol{\beta}) &=& g_2(\boldsymbol{\beta}; \boldsymbol{\mu}^\ast, S_{(m\times m)})~\equiv~t_{4}(\boldsymbol{\beta}\vert \boldsymbol{\mu}^\ast, S),\label{g2approx}
\end{eqnarray}
where the notation Skewed-$t(\alpha\vert\mu, \sigma, a, b)$ represents a density function of a skewed $t$-distribution  at $\alpha$ with location $\mu$, scale $\sigma$, degree of freedom $a+b$, and skewness $a-b$ for any positive constants $a$ and $b$ \citep{jones2003skew}. The article of \cite{jones2003skew} derives the mode of $g_1(\alpha)$ as
\begin{equation}\label{mode}
\mu+\frac{(a-b)\sqrt{a+b}}{\sqrt{(2a+1)(2b+1)}}.
\end{equation} 
The article also show that the tails of the skewed-$t$ density function  follow a power law as $\alpha^{-(2a+1)}$ on the left and $\alpha^{-(2b+1)}$ on the right when $b>a$. It also provides a representation to generate the random variable following their skewed-$t$ distribution as
\begin{equation}
\alpha\sim \mu+\sigma\frac{\sqrt{a+b}(2T-1)}{2\sqrt{T(1-T)}}, ~\textrm{where}~ T\sim \textrm{Beta}(a, b).
\end{equation}
The notation $t_{4}(\boldsymbol{\beta}\vert\boldsymbol{\mu}^\ast, S)$ in Equation \ref{g2approx} indicates a density function of a multivariate $t$-distribution at $\boldsymbol{\beta}$ with 4 degrees of freedom, a location vector $\boldsymbol{\mu}^\ast$, and a $m\times m$ scale matrix $S$ that leads to the variance-covariance matrix $2S$.

@@

\subsection{Model fitting}
The function \code{gbp} creates an S3 object ``gbp'' on which  three generic functions \code{plot}, \code{print}, and \code{summary} are defined. To use these relevant functions, we need to save the outcome of \code{gbp} into an object (see the code below).

@@

The function \code{gbp} contains several optional arguments. The argument \code{Alpha}, whose default is 0.95, sets the confidence level, producing $100\times$\code{Alpha}\% interval estimates of the random effects. For the Gaussian model, setting the argument \code{normal.CI} to \code{TRUE} lets the function \code{gbp} to use a Normal approximation to the unconditional posterior distribution of the random effect \citep{tang2011}. The default of \code{normal.CI} is \code{FALSE} for the skewed Normal approximation \citep{kelly2014advances}.

@@

The S3 object ``gbp'', for example \code{b.output} above, contains detailed outcomes. The function \code{names} will show the list of outcomes (too many to be listed in this article); see the document of \code{gbp} to check the details of the items.
\begin{CodeChunk}
\begin{CodeInput}
R> names(b.output)
R> ?gbp
\end{CodeInput}
\end{CodeChunk}

The S3 object makes use of some functions, \code{print}, \code{summary}, and \code{plot}. The estimation result for all the random effects pops up if we type the ``gbp'' object at the \proglang{R} console, which is the same as the following function \code{print} with its default argument ``\code{sort = TRUE}'' that prints out the estimation result for all the groups in the ascending order of $n$ for the Binomial and Poisson model and the ascending order of standard errors for the Gaussian model. When the argument \code{sort} is \code{FALSE}, the estimation result comes out as the order of data input.
\begin{CodeChunk}
\begin{CodeInput}
R> b.output
R> print(b.output, sort = FALSE)
\end{CodeInput}
\end{CodeChunk}
